---
title: ""
date: ""
author: ""
output:
  md_document:
    variant: markdown_github
---


```{r,echo=FALSE}
library(lattice)
library(knitr)
opts_chunk$set(comment='.', fig.align="center", fig.path="img/README-")

```


[![Build Status master](https://travis-ci.org/metrumresearchgroup/mrgsolve.svg?branch=master)](https://travis-ci.org/metrumresearchgroup/mrgsolve.svg?branch=master)
[![questions](https://www.codewake.com/badges/ask_question.svg)](https://groups.google.com/a/metrumrg.com/forum/#!forum/mrgsolve)



# Resources

  * [Latest release: 0.7.7](https://github.com/metrumresearchgroup/mrgsolve/releases)
  * [Get started with `mrgsolve`](https://github.com/metrumresearchgroup/mrgsolve/wiki/howdoi)
  * Documentation
      * [User Guide](https://mrgsolve.github.io/user_guide)
      * [R docs](https://mrgsolve.github.io/docs)
  * [Installation guide](https://github.com/metrumresearchgroup/mrgsolve/wiki/mrgsolve-installation)
  * [Gallery](https://github.com/metrumresearchgroup/mrgsolve/wiki/gallery)
  * [Extended examples](https://github.com/mrgsolve/examples)
  * [Slide decks](https://github.com/mrgsolve/slides)
  * [Wiki](https://github.com/metrumresearchgroup/mrgsolve/wiki)
     * [Upcoming changes](https://github.com/metrumresearchgroup/mrgsolve/blob/master/NEWS.md)
     * [Windows issues](https://github.com/metrumresearchgroup/mrgsolve/wiki/Windows-issues#issues-with-using-mrgsolve-on-windows-platforms)
  * `mrgsolve` questions and discussion (please use them ... we welcome questions, comments, and issue reports)
     * [Issue tracker](https://github.com/metrumresearchgroup/mrgsolve/issues) (requires GitHub account; ok for questions or issue reports)
     * [Google Group](https://groups.google.com/a/metrumrg.com/forum/#!forum/mrgsolve) (email list or web-based discussion)

<hr>

# Installation

Four primary options exist to install `mrgsolve`:


  * The development version from github via:

```{r,eval=FALSE}
  devtools::install_github("metrumresearchgroup/mrgsolve")
```


  * Latest stable release directly from github via:
  
```{r,eval=FALSE}
  devtools::install_github("metrumresearchgroup/mrgsolve@v0.7.7")
```


  * The latest stable release can be installed with `install.packages` like this:
```{r,eval=FALSE}
  install.packages("mrgsolve",
                   repos=c(options()$repos,"https://mrgsolve.github.io/pkg"),
                   type="source")
```

  * Installation from source the source tar.gz file and all dependencies, [as documented in the wiki](https://github.com/metrumresearchgroup/mrgsolve/wiki/mrgsolve-Installation)



Note that installation from `github` will automatically install all dependencies, such `RcppArmadillo`.



# About
`mrgsolve` facilitates simulation in R from hierarchical,
ordinary differential equation (ODE) based models typically
employed in drug development. The modeler creates a model specification
file consisting of `R` and `C++` code that is parsed, compiled, and dynamically
loaded into the `R` session. Input data are passed in and simulated data are
returned as `R` objects, so disk access is never required during the simulation
cycle after compiling.

## `mrgsolve` is open-source software distributed as a package for `R`

* Installs via `R` package compile and install system
* Easy integration with any other relevant `R` functionality
   * Graphics: `lattice`, `ggplot2`
     * Model estimation: `nls`,`nlme`, `MCMCpack`, `saemix`, others
     * Optimal design: `PFIM`, `PopED`
     * Data summary: `dplyr` and many other functions and packages
     * Interactive model exploration with `shiny`

```{r,message=FALSE}
library(mrgsolve)
library(dplyr)
library(ggplot2)
```



## The model specification file is similar to other non-linear mixed effects modeling software

```{r}
code <- '
$GLOBAL
#define CP (CENT/VC)
#define INH (CP/(IC50+CP))

$SET delta=0.1

$PARAM TVCL=1, TVVC=20, KA = 1.3, KIN=100, KOUT=2, IC50=10

$CMT GUT, CENT, RESP

$MAIN
  double CL = exp(log(TVCL) + ETA(1));
  double VC = exp(log(TVVC) + ETA(2));

  RESP_0 = KIN/KOUT;

$OMEGA 0 0

$ODE
  dxdt_GUT = -KA*GUT;
  dxdt_CENT = KA*GUT - (CL/VC)*CENT;
  dxdt_RESP = KIN*(1-INH) - KOUT*RESP;

$CAPTURE CP
'
```


The model is parsed, compiled, and dynamically loaded into the `R` session

  * Information about the model is saved as an `R` object
  * Important model attributes can be updated in `R` without recompiling


```{r,message=FALSE}
mod <- mcode("demo", code)
```


# Use `mrgsolve` as an interactive simulation tool for model exploration and sensitivity analyses
   * Simulated data are returned as `R` objects
   * Input and output data are kept in memory in the `R` process; writing or reading to disk
   is never necessary (unless results are to be saved for later use).

```{r}
out <- 
  mod %>%
  ev(amt=100, ii=24, addl=2) %>%
  mrgsim(end=120)

out
```



```{r,fig.height=3, fig.width=7}
plot(out, CP+RESP~.)
```

```{r}
out <- 
  mod %>%
  ev(amt=100, ii=24, addl=2) %>%
  Req(CP,RESP) %>%
  knobs(TVVC=c(10,20,40), TVCL=c(0.5,1.5))
```



```{r,fig.height=4,fig.width=9}
plot(out, auto.key=list(columns=3))
```



## Also use `mrgsolve` for efficient, large-scale population simulation

```{r}
mod <- mod %>% omat(cmat(0.1, 0.67, 0.4))
```


## Flexibility with input data sets
  * Data set format that is likely familiar to modeling and simulation scientists
  * No need to include observation records; `mrgsolve` will automatically insert

```{r}
.data <- 
  expand.ev(ID=1:10, amt=c(100,300,1000)) %>%
  mutate(dose=amt)

head(.data)
```


### Input data are passed in as `R` objects
 * Pass many different data sets or implement different designs in the same model code
 without recompiling
 * Control simulation output from `R` to better manage memory

```{r}
out <- 
  mod %>%
  data_set(.data) %>%
  Req(RESP,CP) %>% obsonly %>%
  carry.out(dose) %>%
  mrgsim(end=48, seed=1010)
```

```{r,fig.height=4,fig.width=9}
plot(out, RESP~time|factor(dose), scales="same")
```


## Pass simulated output to your favorite data summary or visualization routines

Summarise with `dplyr`

```{r}
out %>%
  as.tbl %>%
  group_by(dose) %>%
  summarise(rmin = min(RESP), tmim=time[which.min(RESP)])
```


Plot with `ggplot2`

```{r,fig.height=4,fig.width=9,fig.align="center"}
out %>%
  as.tbl %>%
  ggplot(data=.) +
  geom_line(aes(x=time, y=RESP, group=ID, col=factor(dose)))

```
