---
title: ""
date: ""
author: ""
output:
  md_document:
    variant: markdown_github
---


```{r,echo=FALSE}
library(lattice)
library(knitr)
opts_chunk$set(comment='.', fig.align="center", fig.path="img/README-")

```


[![Build Status master](https://travis-ci.org/metrumresearchgroup/mrgsolve.svg?branch=master)](https://travis-ci.org/metrumresearchgroup/mrgsolve.svg?branch=master)
[![questions](https://www.codewake.com/badges/ask_question.svg)](https://groups.google.com/a/metrumrg.com/forum/#!forum/mrgsolve)


# About
`mrgsolve` facilitates simulation in R from hierarchical,
ordinary differential equation (ODE) based models typically
employed in drug development. See the example below.


# Installation
```{r,eval=FALSE}
  devtools::install_github("metrumresearchgroup/mrgsolve")
```
Also see important install-related information [here](https://github.com/metrumresearchgroup/mrgsolve/wiki/mrgsolve-Installation).

# Ask a question
  * [Issue tracker (preferred)](https://github.com/metrumresearchgroup/mrgsolve/issues) (requires GitHub account; ok for questions or issue reports)
  * [Google Group](https://groups.google.com/a/metrumrg.com/forum/#!forum/mrgsolve) (email list or web-based discussion)

# Resources
Please see [mrgsolve.github.io](https://mrgsolve.github.io) for additional resources.

<hr>

# Example

```{r,message=FALSE}
library(mrgsolve)
library(dplyr)
library(ggplot2)
```



## The model specification file is similar to other non-linear mixed effects modeling software

```{r}
code <- '
$GLOBAL
#define CP (CENT/VC)
#define INH (CP/(IC50+CP))

$SET delta=0.1

$PARAM TVCL=1, TVVC=20, KA = 1.3, KIN=100, KOUT=2, IC50=10

$CMT GUT CENT RESP

$MAIN
  double CL = exp(log(TVCL) + ETA(1));
  double VC = exp(log(TVVC) + ETA(2));

  RESP_0 = KIN/KOUT;

$OMEGA 0 0

$ODE
  dxdt_GUT = -KA*GUT;
  dxdt_CENT = KA*GUT - (CL/VC)*CENT;
  dxdt_RESP = KIN*(1-INH) - KOUT*RESP;

$CAPTURE CP
'
```


The model is parsed, compiled, and dynamically loaded into the `R` session

  * Information about the model is saved as an `R` object
  * Important model attributes can be updated in `R` without recompiling


```{r,message=FALSE}
mod <- mcode("demo", code)
```


# Use `mrgsolve` as an interactive simulation tool for model exploration and sensitivity analyses
   * Simulated data are returned as `R` objects
   * Input and output data are kept in memory in the `R` process; writing or reading to disk
   is never necessary (unless results are to be saved for later use).

```{r}
out <- 
  mod %>%
  ev(amt=100, ii=24, addl=2) %>%
  mrgsim(end=120)

out
```



```{r,fig.height=3, fig.width=7}
plot(out, CP+RESP~.)
```

```{r}
out <- 
  mod %>% update(end=48) %>%
  ev(amt=100) %>%
  Req(CP,RESP) %>%
  knobs(TVVC=c(10,20,40), TVCL=c(0.5,1.5))
```



```{r,fig.height=5,fig.width=8}
plot(out)
```



## Also use `mrgsolve` for efficient, large-scale population simulation

```{r}
mod <- mod %>% omat(cmat(0.1, 0.67, 0.4))
```


## Flexibility with input data sets
  * Data set format that is likely familiar to modeling and simulation scientists
  * No need to include observation records; `mrgsolve` will automatically insert

```{r}
.data <- 
  expand.ev(ID=1:10, amt=c(100,300,1000)) %>%
  mutate(dose=amt)

head(.data)
```


### Input data are passed in as `R` objects
 * Pass many different data sets or implement different designs in the same model code
 without recompiling
 * Control simulation output from `R` to better manage memory

```{r}
out <- 
  mod %>%
  data_set(.data) %>%
  Req(RESP,CP) %>% obsonly %>%
  carry_out(dose) %>%
  mrgsim(end=48, seed=1010)
```

```{r,fig.height=4,fig.width=8}
plot(out, RESP~time|factor(dose), scales="same")
```


## Pass simulated output to your favorite data summary or visualization routines

Summarise with `dplyr`

```{r}
out %>%
  as.tbl %>%
  group_by(dose) %>%
  summarise(rmin = min(RESP), tmim=time[which.min(RESP)])
```


Plot with `ggplot2`

```{r,fig.height=4,fig.width=9,fig.align="center"}
out %>%
  as.tbl %>%
  ggplot(data=.) +
  geom_line(aes(x=time, y=RESP, group=ID, col=factor(dose)))

```
