---
pagetitle: "mrgsolve"
date: ""
author: ""
output: github_document
---


```{r,echo=FALSE}
library(lattice)
library(knitr)
opts_chunk$set(comment='.', fig.align="center", fig.path="inst/maintenance/img/README-")
```

# mrgsolve <img align="right" src = "inst/maintenance/img/mrgsolve_sticker_812418_1.png">

[![Build Status](https://travis-ci.org/metrumresearchgroup/mrgsolve.svg?branch=master)](https://travis-ci.org/metrumresearchgroup/mrgsolve)
[![CRAN](http://www.r-pkg.org/badges/version/mrgsolve)](https://cran.r-project.org/package=mrgsolve)
[![License](http://img.shields.io/badge/license-GPL%20%28%3E=%202%29-brightgreen.svg?style=flat)](http://www.gnu.org/licenses/gpl-2.0.html)
[![questions](https://img.shields.io/badge/ask_for-Help-brightgreen.svg)](https://github.com/metrumresearchgroup/mrgsolve/issues)
[![Metrumrg](https://img.shields.io/badge/contact-MetrumRG-brightgreen.svg)](http://metrumrg.com)



mrgsolve is an R package for simulation from hierarchical, ordinary differential
equation (ODE) based models typically employed in drug development.


## Resources
Please see [mrgsolve.github.io](https://mrgsolve.github.io) for additional
resources, including:

- [User Guide](https://mrgsolve.github.io/user_guide)
- [R Documentation](https://mrgsolve.github.io/docs)
- [Vignettes](https://mrgsolve.github.io/vignettes) 
- [Gallery](https://github.com/mrgsolve/gallery) 


## Installation

We recommend staying up to date with the development version
```{r,eval=FALSE}
devtools::install_github("metrumresearchgroup/mrgsolve")
```

Otherwise, install the latest release on CRAN
```{r,eval=FALSE}
install.packages("mrgsolve")
```

__Please__ be sure to see important install-related information 
[here](https://github.com/metrumresearchgroup/mrgsolve/wiki/mrgsolve-Installation).

## Ask a question

We welcome questions about anything: installation, getting your model to 
work, understanding better how mrgsolve works. Please
ask at the [Issue Tracker](https://github.com/metrumresearchgroup/mrgsolve/issues). 
This requires a GitHub account.  But we accept all types of questions there.

