---
title: "Event objects"
author: ""
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    toc: true
    toc_depth: 1
    number_sections: true
    keep_md: true
    fig_width: 5
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{Event objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE, code = readLines("global.R")}
set.seed(23)
```

```{r setup2}
library(mrgsolve)
library(dplyr)
```

# Introduction

Event objects are simple ways to implement PK dosing events into your model 
simulation.  

Let's illustrate event objects with a one-compartment, PK model.  We read this 
model from the `mrgsolve` internal model library.

```{r,message=FALSE}
mod <- modlib("pk1cmt", end = 216, delta = 0.1)
```

# Events

Events are constructed with the `ev()` function

```{r}
e <- ev(amt = 100, ii = 24, addl = 6)
```

This will implement 100 unit doses every 24 hours for a total of 7 doses.  `e` 
has class `ev`, but really it is just a data frame

```{r}
e
```

```{r}
as.data.frame(e)
```

We can implement this series of doses by passing `e` in as the `events` argument 
to `mrgsim`

```{r}
mod %>% mrgsim(events = e) %>% plot(EV1 + CP ~ time)
```


The events can also be implemented with the `ev` constructor along the 
simulation pipeline

```{r}
mod %>%
  ev(amt = 100, ii = 24, addl = 6) %>%
  mrgsim %>% 
  plot(CP ~ time)
```


## Event defaults and expectations

1. `amt` is required
1. `evid = 0` is forbidden
1. Default `time` is 0
1. Default `evid` is 1
1. Default `cmt` is 1

Also by default, `rate`, `ss` and `ii` are 0.

# Combine events

`mrgsolve` has operators defined that allow you to combine events.  Let's first 
define some event objects.

```{r}
e1 <- ev(amt = 500)
e2 <- ev(amt = 250, ii = 24, addl = 4)
e3 <- ev(amt = 500, ii = 24, addl = 0)
e4 <- ev(amt = 250, ii = 24, addl = 4, time = 24)
```


We can combine `e1` and `e3` with a collection operator

```{r}
c(e1, e4)
```

`mrgsolve` also defines a `%then$` operator that lets you execute one event and 
`%then%` a second event
```{r}
e3 %then% e2
```

Notice that `e3` has both `ii` and `addl` defined.  This is required for 
`mrgsolve` to know when to start `e2`.


## Combine to create a data set

We can take several event objects and combine them into a single simulation data 
frame with the `as_data_set` function.

```{r}
e1 <- ev(amt = 100, ii = 24, addl = 6,  ID = 1:5)
e2 <- ev(amt = 50,  ii = 12, addl = 13, ID = 1:3)
e3 <- ev(amt = 200, ii = 24, addl = 2,  ID = 1:2)
```

When combined into a data set, we get 
* N=5 IDs receiving 100 mg Q24h x7 
* N=3 IDs receiving 50  mg Q12h x 14
* N=2 IDs receiving 200 mg Q48h x 3

```{r}
data <- as_data_set(e1, e2, e3)

data
```

To simulate from this data set, we use the `data_set` function.  First, let's 
load a population PK model

```{r,message=FALSE}
mod <- modlib("popex")
```
```{r}
mod %>% data_set(data) %>% mrgsim(end=336) %>% plot("GUT, DV")
```


