---
title: "Get started"
author: ""
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    toc: true
    number_sections: true
    keep_md: true
    fig_width: 5
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE, code = readLines("global.R")}
set.seed(20202)
```

This vignette shows the user how to do their first simulation with mrgsolve.

# Load a model

After loading the package
```{r}
library(dplyr)
library(mrgsolve)
```

we'll use the `modlib()` function to read, parse, compile and load a model 
out of the internal model library.  These models are good to try when learning 
mrgsolve.  

For this first example, we're using a one compartment model (called `pk1` in the 
model library)

```{r}
mod <- modlib("pk1")
```

Once we create the model object (`mod`) we can extract certain information 
about the model.  For example, get an overview 

```{r}
mod
```

Or look at model parameters and their values
```{r}
param(mod)
```

We'll see in the section below how to run a simulation from this model.  But
first, let's create a dosing intervention to use with this model.

# Create an intervention

The simplest way to create an intervention is to use an event object.  This
is just a simple expression of one or more model interventions (most frequently
doses)

```{r}
evnt <- ev(amt = 100, ii = 24, addl = 9)
```

```{r}
evnt
```

We'll see in other vignettes how to create more-complicated events as 
well as how to create larger data sets with many individuals in them.

For now, the event object is just an intervention that we can combine
with a model.

# Simulate

To simulate, use the `mrgsim()` function.
```{r}
out <- mod %>% ev(evnt) %>% mrgsim(end = 480, delta = 0.1)
```

We use a pipe sequence here (`%>%`) to pass the model object (`mod)`
into the event function (`ev()`) where we attach the dosing intervention, 
and that result gets passed into the simulation function.  The result 
of the simulation function is essentially a data frame of simulated values

```{r}
out
```

# Plot

mrgsolve comes with several different functions for processing output.  One 
simple function is a plot method that will generate a plot of what was 
just simulated

```{r}
plot(out)
```
