---
title: "Steady-state"
author: ""
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    toc: true
    toc_depth: 1
    number_sections: true
    keep_md: true
    fig_width: 5
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{Steady-state}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}
source("global.R")
set.seed(2342)
```

# Introduction

When mrgsolve `0.10.0` was released in October, 2019, we made some changes to how
the PK system was advanced to steady state.  This gave better control over the 
process and more-faithful result.  But it also created opportunities for users
to see warning messages that they hadn't seen before.  Further changes were
made in the 0.10.1 release in February, 2020.  

This blog post describes the mechanism by which mrgsolve finds steady state
under this new change set (starting late 2019 to early 2020) and describes 
some of the settings that users can control to their advantage. 

## tl;dr

- pass `ss_n` to put an upper cap on number of doses to give when finding SS
- pass `ss_fixed` to silence warnings when ss isn't met
- set `ss_cmt` (in `[ set ] `) to include or exclude certain compartments for 
  consideration when finding SS
- use `SS_ADVANCE` logical flag in `[ ode ]` to not advance certain compartments
  when mrgsolve is working to find SS

# How does mrgsolve advance the PK system to steady state?

It's important to recognize that `SS` is related to the PK dosing system.  It is
finding the state of the system after an infinite number of doses have been 
administered under a certain regimen.  And this is essentially how mrgsolve
goes about finding steady state. 


## For example

Let's load a PK model

```{r}
library(dplyr)
library(mrgsolve)

mod <- modlib("pk2")
```

And let's imagine a 100 mg QD dosing regimen

```{r}
dose <- ev(amt = 100, ii = 24)
```

And give that dose a large number of times
```{r}
dose <- mutate(dose, total = 1000)
```

Let's just look at trough concentrations and see how the trough develops over 
the dosing period

```{r}
mod <- update(mod, delta = 24, end = 24*20, outvars = "CENT, PERIPH")
```

And simulate that out
```{r}
out <- mrgsim_e(mod, dose, recsort = 3)
```

```{r}
plot(out)
```

Obviously, the trough concentration starts to increase and after a certain period
stops changing

```{r}
tail(out) 
```

Looking at the compartment amount at the last trough compared to the second to 
the last trough, the amounts aren't changing very much.  This is how mrgsolve
finds steady state: it keeps giving the dose at the dosing interval until 
the pre-dose concentration stops changing very much. 

# How close is close enough?

We need some criterion to determine how much change in pre-dose concentration
we'll tolerate an still call it good enough.  mrgsolve uses the local error
estimate for the differential equation solver to determine this.  This is
determined by both `rtol` (the relative tolerance) and `atol` (the absolute
tolerance).  So once the difference between two trough concentrations is less
than A$_{trough}$ * `rtol` + `atol`, then the system is said to be at steady
state.  This calculation is done for every single compartment in the model and
all compartments have to meet this criteria before the system is said to be at
steady state.  So, increasing `rtol` (say from 1e-8 to 1e-3) will give us less
precision in the answer and it will also allow us to call it "good" with respect
to steady state sooner. When using one of the analytical models (one- and
two-compartment models not solved by ODEs), then changing `rtol` will have no
influence on the answer (the answer is known in closed form) but will continue
to influence how easily and quickly steady state is achieved.


# How many doses?

When the volume is larger, it will take more doses to get to steady state and 
more time and work for us to get to that place. Sometimes, simulated 
(many doses) to reach steady state.  The [mrgsim function](https://mrgsolve.github.io/docs/reference/mrgsim.html) has an argument
called `ss_n`.  This is the maximum number of doses that will be given when 
trying to find steady state before mrgsolve gives up, issues a warning and 
keeps going with the problem.  The default value is 500.  So, if the system
gives 500 doses and still can't say that the system is at steady state, it will 
stop, issue a warning and keep going with the problem.  If you were expecting
the system to reach steady state before 500 doses, then it might be good to 
go back and look at the model structure or look at the parameters.

It might be that just a handful of individuals are taking a very long time and 
you're fine with just cutting them off after the 500th dose.  In that case, 
you can invoke the `ss_fixed` argument.  `ss_fixed` is false by default, ensuring
that you will get a warning message if the system fails to reach steady state.
But if you set `ss_fixed` to true, then the system will give **up to** `ss_n` 
doses and stop without warning.  So, here are the possibilities with `ss_n` equal
to 500:

- SS is reached in 10 doses: great! stop after 10 doses and keep going regardless
  of whether `ss_fixed` is true or false
- SS is not reached by the 500th dose and `ss_fixed` is false: keep going after
  issuing a warning that the SS process failed
- SS is not reached by the 500th dose and `ss_fixed` is true: keep going without
  issuing a warning
  

#  What could possibly go wrong?

Well, you might have seen some warnings come up and wondered why.  

## Long half-life

Sometimes it happens that you have simulated parameters with very long half-life
and very long time to steady state, even more than 500 doses.  It might happen
in 1 or 2 out of 3000 simulated individuals and it will still give the warning.
So it might be that you have to set `ss_n` to something reasonable and also set
`ss_fixed` to true so that you don't get the warnings.

## One compartment misbehaves

I have seen this with the dosing compartment when there is a very large 
inter-dose interval.  So the amount gets driven very small, maybe flips sign
and the calculations for steady state just don't work out right to call it 
"good".  Here's what you can do:

In your model, you can now give a vector of compartment to not look at 
for handling steady state:

```{c, eval = FALSE}
[ set ] ss_cmt = "-DEPOT"

[ cmt ] DEPOT CENT PERIPH

[ ode ]
dxdt_DEPOT  = ...;
dxdt_CENT   = ...;
dxdt_PERIPH = ...;
```

The `-DEPOT` says "forget about the DEPOT compartment when running up to 
steady state"; it was giving me numerical problems and I don't really care about 
that one so much either.

## AUC compartment in the model

Let's say you're accumulating stuff in a compartment for AUC calculation.  When 
you have this in the model, you'll never make it to steady state according to 
the definition set out above.  

```{c, eval = FALSE}
[ set ] ss_cmt = "-AUC"

[ cmt ] DEPOT CENT PERIPH AUC

[ ode ]
dxdt_DEPOT  = ...;
dxdt_CENT   = ...;
dxdt_PERIPH = ...;
dxdt_AUC = CENT/V;
```

Rather than trying to figure that out under the hood, mrgsolve just asks you 
to tell it to forget about the `AUC` compartment when determining steady state.

There is another (better) way to handle this. Users have a new variable that 
they can check that evaluates to `true` when mrgsolve is advancing the system
to steady state.  So you might write this:

```{c, eval = FALSE}
[ cmt ] DEPOT CENT PERIPH AUC

[ ode ]
dxdt_DEPOT  = ...;
dxdt_CENT   = ...;
dxdt_PERIPH = ...;
dxdt_AUC = CENT/V;

if(SS_ADVANCE) dxdt_AUC = 0;
```

This will prevent the `AUC` compartment from advancing at all when mrgsolve is 
looking for steady state.  And this compartment won't be a stumbling block for 
the SS determination (as described above). 


## Only consider a single compartment

Rather than excluding the misbehaving compartment, we can also request that 
only one compartment be evaluated for SS

```{c, eval = FALSE}
[ set ] ss_cmt = "CENT"

[ cmt ] DEPOT CENT PERIPH 

[ ode ]
dxdt_DEPOT  = ...;
dxdt_CENT   = ...;
dxdt_PERIPH = ...;
```

This ignores every compartment except for `CENT` when figuring out SS. 

# Summary

- pass `ss_n` to put an upper cap on number of doses to give when finding SS
- pass `ss_fixed` to silence warnings when ss isn't met
- set `ss_cmt` (in `[ set ] `) to include or exclude certain compartments for 
  consideration when finding SS
- use `SS_ADVANCE` logical flag in `[ ode ]` to not advance certain compartments
  when mrgsolve is working to find SS

# Conclusion

I hope this has been helpful to explain steady state concepts in mrgsolve and 
you have better control of this process in your modeling workflow.  I do see 
some additional opportunities potentially coming in the future, like asking
the steady state finder to use a different `rtol` than that which is used 
for solving differential equations.  For the time being, we'll play with this
configuration and see what additional changes would be helpful.

