---
title: Models without compartments (dollar-PRED)
author: ""
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    toc: true
    toc_depth: 1
    number_sections: true
    keep_md: true
    fig_width: 5
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{Models without compartments (dollar-PRED)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
source("global.R")
set.seed(10202)
```

```{r pred_setup}
library(mrgsolve)
```

This vignette introduces a new formal code block for writing models
where there are no compartments.  The block is named after the 
analogous NONMEM block called `$PRED`.  This functionality 
has always been possible with mrgsolve, but only now is there a 
code block dedicated to these models. Also, a relaxed set of 
data set constraints have been put in place when these types of 
models are invoked. 


# Example model

As a most-basic model, we look at the `pred1` model in `modlib()`

```{r}
library(mrgsolve)
mod <- mread_cache("pred1", modlib())
```


The model code is
```{c, eval = FALSE, code = mod@code}

```


This is a random-intercept, random slope linear model.  Like other 
models in mrgsolve, you can write parameters (`$PARAM`), and 
random effects (`$OMEGA`).  But the model is actually written in `$PRED`.  

When mrgsolve finds `$PRED`, it will generate an error if it also finds
`$MAIN`, `$TABLE`, or `$ODE`.  However, the code that gets entered into
`$PRED` would function exactly as if you put it in `$TABLE`.

In the example model, the response is a function of the parameter `B`, 
so we'll generate an input data set with some values of `B`

```{r,message = FALSE}
library(dplyr)
data <- tibble(ID = 1, B = exp(rnorm(100, 0,2)))

head(data)
```

```{r}
out <- mrgsim_d(mod,data,carry.out="B")

plot(out, Y~B)
```


Like other models, we can simulate from a population

```{r}
library(purrr)
set.seed(223)
df <- map_df(1:30, ~ tibble(ID = .x, B = seq(0,30,1)))

head(df)

mod %>% 
  data_set(df) %>% 
  mrgsim(carry.out="B") %>%
  plot(Y ~ B)
```


# PK/PD Model

Here is an implementation of a PK/PD model using `$PRED`

In this model

- Calculate `CL` as a function of `WT` and a random effect
- Derive `AUC` from `CL` and `DOSE` 
- The response (`Y`) is a calculated from `AUC` and the Emax 
  model parameters

```{r}
code <- '
$PARAM TVCL = 1, WT = 70, AUC50 = 20, DOSE = 100, E0 = 35, EMAX = 2.4

$OMEGA 1

$SIGMA 100

$PRED
double CL = TVCL*pow(WT/70,0.75)*exp(ETA(1));
capture AUC = DOSE/CL;
capture Y = E0*(1+EMAX*AUC/(AUC50+AUC))+EPS(1);
'
```


```{r}
mod <- mcode_cache("pkpd", code)
```

To simulate, look at 50 subjects at each of 5 doses
```{r}
data <- 
  expand.idata(DOSE = c(30,50,80,110,200),ID = 1:50) %>% 
  mutate(WT = exp(rnorm(n(),log(80),1)))

head(data)
```

```{r}
out <- mrgsim_d(mod,data,carry.out="WT,DOSE") %>% as.data.frame()

head(out)
```


Plot the response (`Y`) versus `AUC`, colored by dose
```{r}
library(ggplot2)

ggplot(out, aes(AUC,Y,col =factor(DOSE))) + 
  geom_point() + 
  scale_x_continuous(trans = "log", breaks = 10^seq(-4,4)) + 
  geom_smooth(aes(AUC,Y),se = FALSE,col="darkgrey") + theme_bw() + 
  scale_color_brewer(palette = "Set2", name = "") + 
  theme(legend.position = "top")
```

