---
title: Set initial conditions
author: ""
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    toc: true
    toc_depth: 1
    number_sections: true
    keep_md: true
    fig_width: 5
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{Set initial conditions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE}
source("global.R")
set.seed(1010)
```


# Short answer
There are two commonly-used ways to set initial conditions: in `$MAIN` and in
the initial condition list.

## Set initials in `$MAIN`

For a compartment called `CMT`, there is a variable available to you called
`CMT_0` that you can use to set the initial condition of that compartment in
`$MAIN`.  For example:

```{r}
code <- '
$PARAM KIN = 200, KOUT = 50

$CMT RESP

$MAIN
RESP_0 = KIN/KOUT;
'
```
This is the most commonly-used way to set initial conditions: the initial condition for the `RESP` compartment is set equal to `KIN` divided by `KOUT`.  If you had a parameter called `BASE`, you could also write `RESP_0 = BASE;`.  In these examples, we're using data items from `$PARAM`.  But the initial condition could be set to any numeric value in the model, including individual parameters derived from parameters, covariates, and random effects.  Note that you should never declare `RESP_0` (e.g. `double RESP_0`): it just appears for you to use.

## Set initials in the `init` list
You can also set initial conditions in the initials list.  Most commonly, this means declaring compartments with `$INIT` rather than `$CMT`.   For example

```{r}
code <- '
$INIT RESP = 4
'
```

This method gets us the same result as the previous example, however the initial condition now is not a derived value, but it is coded as a number.  Alternatively, you could declare a compartment via `$CMT` and update later (see next).

We can update this value later like this
```{r}
library(mrgsolve) 

mod <- mcode("init_up", code)

init(mod)

init(mod, RESP=8) %>% init
```

This method is commonly used to set initial conditions in large QSP models where the compartment starts out as some known or assumed steady state value.

## Don't use initial conditions as a dosing mechanism
Using an initial condition to put a starting dose in a compartment is not recommended.  Always use a dosing event for that.




# Long answer

The following is from a wiki post I did on the topic.  It's pedantic.  But hopefully helpful to learn what `mrgsolve` is doing for those who want to know.

* `mrgsolve` keeps a base list of compartments and initial conditions that you can update __either__ from `R` or from inside the model specification
    * When you use `$CMT`, the value in that base list is assumed to be 0 for every compartment
    * `mrgsolve` will by default use the values in that base list when starting the problem
    * When only the base list is available, every individual will get the same initial condition
* You can __override__ this base list by including code in `$MAIN` to set the initial condition
    * Most often, you do this so that the initial is calculated as a function of a parameter
    * For example, `$MAIN RESP_0 = KIN/KOUT;` when `KIN` and `KOUT` have some value in `$PARAM`
    * This code in `$MAIN` overwrites the value in the base list for the current `ID`
* For typical PK/PD type models, we most frequently initialize in `$MAIN`
    * This is equivalent to what you might do in your NONMEM model
* For larger systems models, we often just set the initial value via the base list
        

## Make a model only to examine `init` behavior

Note: `IFLAG` is my invention only for this demo.  The demo
is always responsible for setting and interpreting the value (it is not 
reserved in any way and `mrgsolve` does not control the value).

For this demo

* Compartment `A` initial condition defaults to 0
* Compartment `A` initial condition will get set to `BASE` __only__ if `IFLAG  > 0`
* Compartment `A` always stays at the initial condition (the system doesn't advance)

```{r}
code <- '
$PARAM BASE=200, IFLAG = 0

$CMT A

$MAIN
if(IFLAG > 0) A_0 = BASE;

$ODE dxdt_A = 0;
'
```


```{r}
mod <- mcode("init_long",code)
dplot <- function(x) plot(x,ylim=c(0,400))
```

__Check the initial condition__
```{r}
init(mod)
```

Note:

* We used `$CMT` in the model spec; that implies that the base initial condition for `A` is set to 0
* In this chunk, the code in `$MAIN` doesn't get run because `IFLAG` is 0
* So, if we don't update something in `$MAIN` the initial condition is as we set it in the base list

```{r}
mod %>% mrgsim %>% dplot
```

__Next, we update the base initial condition for `A` to 100__

Note:

* The code in `$MAIN` still doesn't get run because `IFLAG` is 0

```{r}
mod %>% init(A = 100) %>% mrgsim %>% dplot
```

__Now, turn on `IFLAG`__

Note:

* Now, that code in `$MAIN` gets run
* `A_0` is set to the value of `BASE`

```{r}
mod %>% param(IFLAG=1) %>% mrgsim %>% dplot

mod %>% param(IFLAG=1, BASE=300) %>% mrgsim %>% dplot
```


## Example PK/PD model with initial condition

Just to be clear, there is no need to set any sort of flag to set the initial condition.


```{r}
code <- '
$PARAM AUC=0, AUC50 = 75, KIN=200, KOUT=5

$CMT RESP

$MAIN 
RESP_0 = KIN/KOUT;

$ODE
dxdt_RESP = KIN*(1-AUC/(AUC50+AUC)) - KOUT*RESP;
'
```


```{r}
mod <- mcode("init_long2", code)
```


The initial condition is set to 40 per the values of `KIN` and `KOUT`

```{r}
mod %>% mrgsim %>% plot
```

Even when we change `RESP_0` in `R`, the calculation in `$MAIN` gets the final say

```{r}
mod %>% init(RESP=1E9) %>% mrgsim
```


## Calling `init` will let you check to see what is going on

* It's a good idea to get in the habit of doing this when things aren't clear
* `init`  first takes the base initial condition list, then calls `$MAIN` and does any calculation 
you have in there; so the result is the calculated initials

```{r}
init(mod)

mod %>% param(KIN=100) %>% init
```


# Set initial conditions via `idata`

Go back to house model

```{r}
mod <- mrgsolve:::house()
init(mod)
```

Notes

* In `idata` (only), include a column with `CMT_0` (like you'd 
do in `$MAIN`).
* When each ID is simulated, the `idata` value will override the 
base initial list for that subject.  
* But note that if 
`CMT_0` is set in `$MAIN`, that will override the `idata` update. 

```{r}
idata <- expand.idata(CENT_0 = seq(0,25,1))

idata %>% head

out <- 
  mod %>% 
  idata_set(idata) %>% 
  mrgsim(end=40)
```


```{r}
plot(out, CENT~.)

```



