---
title: "Time after dose"
author: ""
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    toc: true
    toc_depth: 1
    number_sections: true
    keep_md: true
    fig_width: 5
    fig_height: 4
vignette: >
  %\VignetteIndexEntry{Time after dose}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
source("global.R")
set.seed(10072)
```

# Introduction

`TAD` stands for time after dose.  In mrgsolve, any time `TAD` is calculated
automatically, a dose is a record in the input data set where `EVID` is set to 
either 1 or 4. 

# TAD calculated as runtime argument

For a while, you've been able to get time after dose in your simulated output

```{r, message = FALSE, warning=FALSE}
library(mrgsolve)
library(dplyr)

mod <- modlib("pk1", req = "")

mod %>% 
  ev(amt = 100, ii = 4, addl = 3, time = 2) %>%
  mrgsim(tad = TRUE, add = c(3.888,5.91)) %>% 
  as_tibble() %>% 
  head(n = 10)
  
```

This is convenient because you can choose the output at run time, 
we give you the negative numbers prior to the first dose etc.  This 
sort of calculation is possible because we let mrgsolve know 
ahead of time that we want this calculation done and mrgsolve
makes an extra pass through the records to find when is the 
first dose for each individual.

# TAD calculated in the model
Sometimes you would like to work with time after dose in
your model.  This isn't super-complicated to do but does 
require some programming and setup and all of that. As of 
mrgsolve 0.9.1, there is a special function to do these 
calculations for you.

```{r,message=FALSE}
mod <- mread("time-after-dose.txt",req = "")
```

```{r, echo=FALSE}
x <- readLines("time-after-dose.txt",warn=FALSE) %>% modelparse()
```

Looking at the `[MAIN]` block:
```{c,eval=FALSE, code = x$MAIN}

```

When you call `self.tad()` you will get the time after dose.  It's
important that this gets called on every record ... specifically every dosing 
record.  It will not work properly if it is not called every record and there 
is no check at this time to make sure you follow that rule.  So please follow 
the rule.

To see an example:

```{c, eval=FALSE, code = mod@code}
```

```{r}
mod %>% 
  ev(amt = 100, ii = 4, addl = 3, time = 2) %>%
  mrgsim(tad = TRUE, add = c(3.888, 5.91)) %>% 
  as_tibble() %>% 
  head(n = 10)
```

You will notice two differences between `tad` (output requested at 
run time) and `tadose` (values calculated in the model itself) in the 
output listing above:  

1. `tadose` is -1 before the first dose
1. specifically, `tadose` is -1 at the 2 hour observation record
   that occurs that the same time as the dose, but happens before 
   the dose in record order

The main point if this is that you can easily obtain time after 
dose in the problem (model) itself to use as you program the model
and also output the number into the simulated output.

# Time after dose in specific compartment

Version 0.10.7 adds a new plugin with the ability to calculate time after dose 
in any compartment. 

We write a model using the `tad` plugin to track time after dose in 
compartments `one` and `two`.  We create `tadose` objects to track 
this and we can call the `tad()` method on these objects, passing 
in the `self` data item.

```{r}
code <- '

[plugin] tad

[ global ] 
mrg::tadose tad_cmt_1(1); 
mrg::tadose tad_cmt_2(2);

[ pkmodel ] cmt = "GUT,CENT", depot = TRUE

[ param ] CL = 1, V = 20, KA = 1

[ main ] 
capture tad1 = tad_cmt_1.tad(self); 
capture tad2 = tad_cmt_2.tad(self);

'
  
mod <- mcode("tad", code, soloc = '.')


data <- c(
  ev(amt = 100, cmt = 1, time = 1), 
  ev(amt = 200, cmt = 2, time = 3)
)

mrgsim(mod, data)

```

Note that time after dose is -1 until a dose is administered. 

Recall also that time after dose can be calculated more simply if there is 
only one dose type by passing the `tad` argument:

```{r}
data1 <- filter(data, cmt ==1) %>% mutate(time = 3)
mod %>% mrgsim(data1, tad = TRUE)
```

This is a little nicer because it will fill in negative `tad` values for you.
 
