---
title: ""
output: 
  pdf_document: 
    number_sections: true
---

```{r,include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
```

# Validation Testing: mrgsolve v0.10.0

## Scope

The purpose of this Validation Testing document is to define the conditions for
test execution. All tests are specified and linked to release candidate user
stories as numbered issues in the Requirements Specification-Validation Plan
document.

----------------

```{r, message=FALSE, warning = FALSE}
library(dplyr)
library(readr)
library(testthat)
```


## Test locations

Tests are in three locations

1. `tests/testthat` 
1. `inst/maintenance/unit` 
1. `inst/validation`


## Test candidate

Showing md5 checksum
```{r}
tarball <- "../../mrgsolve_0.10.0.tar.gz"
tools::md5sum(tarball)
```


```{r,output="hide", message=FALSE,warning=FALSE}
install.packages(tarball, repos=NULL)
```


```{r}
run_test <- function(dir,d) {
 cwd <- getwd()
 on.exit(setwd(cwd))
 setwd("../../")
 x <- test_dir(dir, reporter="check") %>% as_tibble
 x$result <- NULL
 x$location <- dir
 x$date <- d
 x
}
```

```{r}
d  <- Sys.time()
```

## Run tests

Tests specified in the Requirements Specification-Validation Plan document are
executed using the `run_test()` function, which calls `testthat::test_dir`.

### tests/testthat

```{r, echo=FALSE}
y <- run_test("tests/testthat",d) 
```


### inst/maintenance/unit

```{r, echo=FALSE}
z <- run_test("inst/maintenance/unit/",d)
```


### inst/validation

```{r, echo=FALSE}
x <- run_test("inst/validation",d)
```

```{r}
df <- bind_rows(x,y,z)
```

## Comprehensive summary

Summarizes the number of contexts, tests, and expectations and counts number
of tests and test failures.

```{r}
summarise(
        df,
        contexts = n_distinct(context),
        tests = n_distinct(test), 
        number = sum(nb), 
        failed = sum(failed)
)
```

```{r}
readr::write_csv(path="all_tests.csv", df)
```

## Session

Testing session information is captured.

```{r}
devtools::session_info()
```
